\name: Combining Bottom-up and top-down approaches through graph learning over interaction networks for drug-target-interaction prediction

### 2021-03-18 paper/main.tex
.

As we want to learn from the similarity of drug side effects and
protein phenotypes, we use a deep half-twin neural network with
a contrastive loss using cosine similarity.
A half-twin neural network aims to learn a similarity between two
embeddings.

% learning a high-dimensional embedding emphasizing this identity by
% forcing a similarity between these embeddings. On the other hand we
% built a deep neural network for the molecular structure based
% features, also benefiting from the siamese network
% architecture. Computing the similarity between two representations,
% allows for a variety of different methods.  
% However, we decided for
% the cosine similarity measure being invariant to scaling.

Therefore, the precomputed embeddings are run through a regular,
neural learnable feature transformation (LFT) network, which also
reduces the eventual representation size for drugs and proteins
separately. An example structure for both types of features can be
found in Figure~\ref{fig:HalfTwinNetwork}.

While a regular deep neural network, denoted by LFT, for feature space reduction is not particularly novel, we emphasize the versatility of this approach, as both ontology and molecular feature for both drugs and proteins are reduced to similar dimensionality. This allows for a high amount of modularity and different experimental setups by plugging different kinds of features into the model. Additionally, these pretrained features can be used for a variety of other tasks. Additionally, the ontology LFT can be reused for a variety of DL2vec based features with respect to other ontologies and hypotheses. We hereby followed the results of \textit{DL2vec}, indicating that utilizing the activation function $\sigma := \mathrm{LeakyReLU}$ leads to performance increase.

>>>>

As we want to learn from the similarity of drug side effects and
protein phenotypes, we use a deep half-twin neural network with a
contrastive loss using cosine similarity.  A half-twin neural network
aims to learn a similarity between two embeddings of the same
dimension.  The precomputed embeddings are first processed by a
learnable feature transformation (LFT) network, a fully connected
neural network layer which takes as input an embedding and outputs a
representation of a particular size. An example structure for both
types of features can be found in Figure~\ref{fig:HalfTwinNetwork}. We
use this LFT to reduce the eventual representation size for drugs and
proteins separately.
% While a regular deep neural network, denoted by LFT, for feature space reduction is not particularly novel, we
The use of an LFT enables flexible experimentation as both ontology
and molecular feature for both drugs and proteins are reduced to
similar dimensionality for varying sizes of inputs; this allows for a
high amount of modularity across different experimental setups by
adding different kinds of features into the model. Additionally, the
generated features can be used for a variety of other tasks; for
example, the ontology LFT can be reused for a variety of DL2vec-based
features with respect to other ontologies and hypotheses. We follow
the results of \textit{DL2vec} showing that the activation function
$\sigma := \mathrm{LeakyReLU}$ leads to improved performance compared
to other activation functions.
